```{r fulfill package requirements, echo=FALSE}
# install package for interfacing with Cytoscape
if(!require("BiocManager")) {
  install.packages("BiocManager", repos = "https://cloud.r-project.org")
  BiocManager::install("RCy3", force=TRUE)
}

# install mappeR package
if(!require("mappeR")) {
  install.packages("mappeR")
}

# load required packages
library(mappeR)
library(RCy3)
library(readxl)
library(stringdist)
library(usedist)
```

First, we import our aptamer data.
```{r import data}
# import from excel file
selex_data = na.omit(as.data.frame(read_xlsx("data/selex_data.xlsx", sheet = "Data")))

# individual observations are aptamers
rownames(selex_data) = selex_data$Name
```

Next, we generate a distance matrix for our aptamers using the Levenshtein distance between sequences.
```{r create distance matrix}
# define a function to get the edit distance between two aptamers
get_edit_distance <- function(apt1, apt2) {
  return(stringdist(apt1["Variable"], apt2["Variable"], method = "lv"))
}

# generate distance matrix for aptamers using edit distance
edit_dists = dist_make(as.matrix(selex_data), get_edit_distance)
```

Now we can create a ballmapper graph from this distance matrix, using some value of $\varepsilon$:
```{r create ballmapper object, echo=FALSE}
# create ball mapper object with edit distances and specified epsilon value
eps = 3
aptamerballs = create_ball_mapper_object(selex_data, edit_dists, eps)
```

In the ballmapper graph, vertices represent collections of aptamers which are close in Levenshtein distance; we can calcuate stats from the SELEX process for each collection.
```{r calculate statistics}
# get aptamers per vertex of the ballmapper graph
balled_aptamers = lapply(aptamerballs[[1]]$data, function(x) selex_data[unlist(strsplit(x, ", ")),])

# calculate median round amounts for the aptamers in each vertex
rounds = c("RP10M 0A", "RP10M 0B", "RP10M 0C", "RP10M 1A", "RP10M 1B", "RP10M 1C", "RP10M 2A", "RP10M 2B", "RP10M 2C", "RP10M 3A", "RP10M 3B", "RP10M 3C", "RP10M 4", "RP10M 5", "RP10M 6", "RP10M 7", "RP10M 8", "RP10M 9")

round_medians = sapply(balled_aptamers, function(x) sapply(rounds, function(r) max(selex_data[x$Name, r])))

# calculate median human affinity values aptamers per vertex
median_human_affinity = sapply(balled_aptamers, function(x) median(selex_data[x$Name, "ASSET hVSMC:hEC"]))
mean_human_affinity = sapply(balled_aptamers, function(x) mean(selex_data[x$Name, "ASSET hVSMC:hEC"]))
max_human_affinity = sapply(balled_aptamers, function(x) max(selex_data[x$Name, "ASSET hVSMC:hEC"]))

# calculate median mouse affinity values of aptamers per vertex
median_mouse_affinity = sapply(balled_aptamers, function(x) median(selex_data[x$Name, "ASSET mVSMC:mEC"]))
mean_mouse_affinity = sapply(balled_aptamers, function(x) mean(selex_data[x$Name, "ASSET mVSMC:mEC"]))
max_mouse_affinity = sapply(balled_aptamers, function(x) max(selex_data[x$Name, "ASSET mVSMC:mEC"]))
```

Now we populate our mapper object with the stats we calculated...
```{r record statistics}
# add affinity info to mapper graph data
aptamerballs[[1]][rounds] = t(round_medians)
aptamerballs[[1]]$median_human_affinity = median_human_affinity
aptamerballs[[1]]$max_human_affinity = max_human_affinity
aptamerballs[[1]]$mean_human_affinity = mean_human_affinity
aptamerballs[[1]]$median_mouse_affinity = median_mouse_affinity
aptamerballs[[1]]$max_mouse_affinity = max_mouse_affinity
aptamerballs[[1]]$mean_mouse_affinity = mean_mouse_affinity
```

...and pass the whole dataset to Cytoscape for visualization.
```{r pass to cytoscape}
# send mapper data to cytoscape; cytoscape must be open before running this chunk
cymapper(aptamerballs, is_ballmapper = TRUE)
```







